### 2020-04-24 Silvers_cfmip_CRE_SinglePanel.tex
addition of three tex files that are used to generate multipanel figures

\documentclass[varwidth, border=10pt]{standalone}

\usepackage{subcaption}
\usepackage[labelformat=parens,labelsep=quad,skip=3pt]{caption}
\usepackage{graphicx}

\begin{document}

\begin{figure}
  \centering
  \begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\linewidth]{cfmipfigs/CRE_net_ceres_cfmipExps.eps}
\end{subfigure}
\begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\linewidth]{cfmipfigs/CRE_SWonly_feedback_amip.eps}
\end{subfigure}
\end{figure}

\end{document}