### 2020-04-24 Silvers_cfmip_MISR_cloudtable.tex

\documentclass[varwidth, border=10pt]{standalone}

%\usepackage{subcaption}
\usepackage{graphicx}
%\usepackage[labelformat=parens,labelsep=quad,skip=3pt]{caption}
\usepackage{subfig}

\begin{document}

\begin{figure}
  \centering
  \subfloat{
  \includegraphics[width=0.2\linewidth]{cfmipfigs/cloud_table_misr_am4_glbmn.eps}}
%\end{subfigure}
  \subfloat{
  \includegraphics[width=0.2\linewidth]{cfmipfigs/cloud_table_misr_am4_sh_30t60.eps}}
  \subfloat{
  \includegraphics[width=0.2\linewidth]{cfmipfigs/cloud_table_misr_am4_pm30.eps}}
\subfloat{
  \includegraphics[width=0.2\linewidth]{cfmipfigs/cloud_table_misr_am4_nh_30t60.eps}}
\end{figure}

%\begin{figure}
%  \centering
%  \begin{subfigure}{0.48\textwidth}
%  \includegraphics[width=\linewidth]{cfmipfigs/cloud_table_misr_am4_glbmn.eps}
%\end{subfigure}
%\begin{subfigure}{0.48\textwidth}
%  \includegraphics[width=\linewidth]{cfmipfigs/cloud_table_misr_am4_sh_30t60.eps}
%\end{subfigure}
%\end{figure}

\end{document}