### 2020-04-24 Silvers_cfmip_SinglePanelFig.tex

\documentclass[varwidth, border=10pt]{standalone}

\usepackage{subcaption}
\usepackage[labelformat=parens,labelsep=quad,skip=3pt]{caption}
\usepackage{graphicx}

\begin{document}

\begin{figure}
  \centering
  \begin{subfigure}{0.48\textwidth}
    \includegraphics[width=\linewidth]{/Users/silvers/Research/cfmip_paper/cf_warming.eps}
\end{subfigure}
\begin{subfigure}{0.48\textwidth}
  \includegraphics[width=\linewidth]{/Users/silvers/Research/cfmip_paper/cf_warming_amipbias.eps}
\end{subfigure}
\end{figure}

\end{document}