### 2020-05-26 Silvers_cfmip_extraTable.tex
extraTable with new feedback numbers from David Paynter

\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

\draftfalse

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}

\begin{document}

We calculate feedbacks for the coupled experiments, and amip, as follows.     The difference in surface temperature is calculated 
relative to the piControl experiment for $T_1$ and as a difference from the first 20 years of the amip experiment for $T_2$.   
The radiative imbalance at TOA, $N$ is calculated over different periods depending on the experiment.  In all cases
the years used for $N$ and $T$ are taken from the same years.  $N_1$ corresponds to years 101-150, $N_2$ corresponds
to years 1996-2015.

\begin{equation}
  \alpha_{4xCO_2}=-\frac{F_{4xCO_2}-N_1}{T_1}
\end{equation}
\begin{equation}
  \alpha_{1\%CO_2}=-\frac{F_{1\%CO_2}-N_1}{T_1}
\end{equation}
\begin{equation}
  \alpha_{hist}=-\frac{F_{hist}-N_2}{T_1}
\end{equation}
\begin{equation}
  \alpha_{amip}=-\frac{F_{amip}-N_2}{T_2}
\end{equation}

In the above equations $F_{4xCO_2}=N_{4xCO2}-N_{PiC}$; $F_{hist}=N_{hist(1996-2014)}-N_{PiC}$; 
$F_{1pctCO_2}=0.79F_{4xCO2}$.

\begin{table}
\begin{center}
\caption{Global mean radiative feedbacks ($\rm{W\, m^{-2} K^{-1}}$).  The cloud radiative effect (CRE) is computed as
 the clear sky minus all sky TOA flux. The values for the Abrupt 4xCO2 experiment
have been estimated from a linear fit of the years... }
    \begin{tabular}{*{5}{c}}
    \hline
    \hline
 Feedback   & Abrupt 4xCO2 & 1pct.   & Historical    &  Amip   \\ \hline
    Net          &   -1.11      &  -1.22          &  -1.83         & -1.77               \\ 
    \\
    Net CRE   & -0.06       &  -0.26          & -0.45         & -0.46           \\  
    \\
    LW CRE   & 0.14        &  0.10            & -0.14           & -0.03          \\  
    \\
    SW CRE  & -0.20       &  -0.26            & -0.31          & -0.43         \\  
    \\
    LW CLR   & -1.70       &  -1.71            & -1.94           & -2.25        \\  
    \\
    SW CLR  & 0.65        & 0.65              & 0.56             & 0.91          \\  
    \\
    $\Delta \rm{T}_s$        & 3.99        &  ??                & 0.8              & 0.76          \\  \hline

    \end{tabular}\par
    %\bigskip 
    \label{tab:lambda}
\end{center}
\end{table}

\end{document}