### 2020-06-20 Silvers_cfmip_CRE_SinglePanel.tex
don't have a clue

\documentclass[varwidth, border=10pt]{standalone}

\usepackage{caption}
\usepackage{subcaption}
\usepackage[labelformat=parens,labelsep=quad,skip=3pt]{caption}
\usepackage{graphicx}

\begin{document}
% for some reason, including a blank line between the subfigure blocks leads to a vertically aligned figure...

\begin{figure}
  \centering
  %\hspace{\fill}
  \begin{subfigure}{0.31\textwidth}
  \centering
  \includegraphics[width=\linewidth]{cfmipfigs/CRE_net_ceres_cfmipExps.eps}
\end{subfigure}
 %\hspace{\fill}
%\vspace{2pt}

 %\hspace{\fill}
\begin{subfigure}{0.31\textwidth}
\centering
  \includegraphics[width=\linewidth]{cfmipfigs/CRE_feedback_fromPiControl_Norm.eps}
\end{subfigure}
 %\hspace{\fill}
%\vspace{2pt}

 %\hspace{\fill}
\begin{subfigure}{0.31\textwidth}
\centering
  \includegraphics[width=\linewidth]{cfmipfigs/CRE_SWonly_feedback_amip.eps}
\end{subfigure}
 %\hspace{\fill}
\end{figure}

\end{document}